LME-MPM applied to quasi-brittle fracture.

### 2020-04-25 2020_EFM_MPM_Eigensoftening.tex
Cambios 24-4-2020

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}

The simulation of fracture propagation in a more accurate and
effective way can be considered as one of the original drivers for
developing novel spatial discretization methods such as meshfree
methods like the material point method (MPM). Presence of cracks are a
violation of the continuity requirement of the finite element method
(FEM). To overcome it, numerous of numerical artifacts has been
proposed with the aim of reproduce such a complex behaviour. This
techniques vary from employing cohesive approaches
\cite{Barenblatt,Hilleborg_1976}, by adaptively inserting cohesive
elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid elements
boundaries, or handling arbitrary cracks paths by level set
representation of the fracture surface \cite{Belytschko_03}.

The severe
limitations of the mesh-dependent methods to simulate fracture
process, have motivated the development of mesh-free technique to
reproduce crack propagation such the element-free Galerkin method
(EFGM) \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013},
the smoothed particle hydrodynamics (SPH) \cite{Wang_2020,Wang_2019},
the optimal transportation meshfree method (OTM)
\cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or peridynamics
\cite{HA_2011,RABCZUK_2017}.

By contrast, MPM does not suffer from the above difficulty since the
continuity requirement is less restrictive. Fracture can be described
in two ways. One is to remove the restriction of the single-valued
velocity field close to the crack by using two or more sets of nodes
\cite{Nairn_2006}. This method is based in to assign different labels
to distinguish if the material points and nodes are in the same side
of the crack or not. Under this approach the crack surface is
described with line segments in 2D and triangle patches in 3D
cases. This method was named as ``CRAcks with Material Points
(CRAMP)'' \cite{Nairn_2003}. In this method, the criteria for crack
propagation is based on such parameters as energy release rate
analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. The other approach is to use failed material
points to describe the crack evolution. In this method, the formation
of failed points describes the nucleation of cracks, and thereafter its propagation and
branching. Consequently, the position of the crack does not need to be
explicitly stated. This represents a significant advantage over the
``CRAMP''. Under this approach, the prediction the failure evolution
employing a decohesion model has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} or
Schreyer {\it et al.}\cite{Schreyer_2002}. And has been successfully
employed to simulate the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
(2004)\nocite{Sulsky_2004}.

\cite{Belytschko_04,Schmidt_2009,Pandolfi_2012}

The erosion of the material point means that each material point can
be either intact or be completely failed or eroded and has no load
bearing capacity, Even thought the method has been successfully
applied to dynamic fragmentation of metals, quantitative validations,
such as stress or strain levels near or at the crack set compared to
experimental measures are still lacking. Having in mind that the
eigenerosion approach was intended for perfectly brittle fracture, an
eigensoftening concept is developed for quasi-brittle materials

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The meshfree methodology}
\label{sec:2}

The aim of this section is to provide an overview of the special
techniques employed to face the fracture problem under the MPM
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the explicit-predictor algorithm for the
MPM will be exposed, next the local \textit{max-ent} approximants are introduced in
\ref{sec:2.2} as an accurate alternative technique to interpolate data
between particles and nodes, and finally the fracture algorithms based
in the eigendeformations are presented in \ref{sec:2.3}.

\subsection{The Material Point Method}
\label{sec:2.1}

The MPM~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method
where particles carries on all the physical information and a set of
background nodes is employed to compute the equilibrium
equation. Since the MPM possesses the advantages of both Lagrangian
and Eulerian descriptions, no element distortion exists in the MPM,
therefore it is an appropriate and efficient method in solving
problems involving extremely large deformation and moving
discontinuities such fracture evolution. For the spatial
discretization, two sets of points are introduced in 
the MPM. First, the nodes. This points are considered fixed in the
space and are in charge of computing all the kinematic fields such
forces $f_I$, accelerations $a_I$ and velocities $v_I$. And second the
material points or particles. They are in charge of the discretization
of the continuum, and store the local state ($\sigma_p,
\varepsilon_p$). Without loosing generality, the MPM algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation thorough a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration
scheme is adopted. The purpose of this choice is motivated due its proved
robustness and stability in numerical calculations. In the first stage
(i), the nodal velocity predictor is computed following
\eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocity are
obtained, the essential boundary conditions are imposed. And in the
following, a Eulerian phase (ii) is computed in the set of nodes in a
FEM-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are computed thorough the
equilibrium equation. Next the nodal velocities are corrected in a
\textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}
\end{equation}
Finally updated the particles are advected in the Lagrangian stage (iii) using nodal values as,
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k} 
\end{align}
The complete pseudo-algorithm it is summarized in \ref{sec:expl-pred-corr}.

@@

\sum_I N_I |\vec{x}_p - \vec{x}_I |^2$ a 
magnitude of the shape function width. The regularization o
\textit{thermalization} parameter between the two criterion $\beta$
has Pareto optimal values in the range $(0,\infty)$. The unique
solution of the local max-ent problem (LME)$_\beta$ is

@@

the limit values of the influence radius and is used thereafter to

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
Presence of cracks are a violation of the continuity requirement of
the finite element method (FEM). To overcome it, numerous of numerical
artifacts has been proposed with the aim of reproducing  such a
complex behaviour. This techniques vary from employing cohesive
approaches \cite{Barenblatt,Hilleborg_1976}, by adaptively inserting
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries, or handling arbitrary cracks paths by level set
representation of the fracture surface \cite{Belytschko_03}.

The simulation of fracture propagation in a more accurate and
effective way can be considered as one of the original drivers for
developing novel spatial discretization methods such as meshfree
techniques. Some examples of it are the material
point method (MPM) \cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
element-free Galerkin method (EFGM) \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013},
the smoothed particle hydrodynamics (SPH) \cite{Wang_2020,Wang_2019},
the optimal transportation meshfree method (OTM)
\cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or peridynamics
\cite{HA_2011,RABCZUK_2017} among others.

Fracture can be described in two ways. One is to remove the
restriction of the single-valued velocity field close to the crack by
using two or more sets of nodes \cite{Nairn_2006}. This method is
based in to assign different labels to distinguish if the material
points and nodes are in the same side of the crack or not. Under this
approach the crack surface is described with line segments in 2D and
triangle patches in 3D cases. This method was named as ``CRAcks with
Material Points (CRAMP)'' \cite{Nairn_2003}. In this method, the
criteria for crack propagation is based on such parameters as energy
release rate analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the
stress intensity factor or the J-integral discussed by Guo \& Nairn 
(2004)\cite{Nairn_2004}. The other approach is to use failed material
points to describe the crack evolution. In this method, the formation
of failed points describes the nucleation of cracks, and thereafter
its propagation and branching. Consequently, the position of the crack
does not need to be explicitly stated. This represents a significant
advantage over the ``CRAMP''. Under this approach, the prediction the
failure evolution employing a decohesion model has been discussed by
Chen {\it et al.}\cite{Zhenmao_2005} or Schreyer {\it et
  al.}\cite{Schreyer_2002}. And has been successfully employed to
simulate the fracture of brittle materials by Chen {\it et al.}
\cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}. Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
they approximate the crack set by means of eigen-deformations, wich
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle facture
was developed by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. In this technique the erosion of
the material point means that each material point can be either intact
or be completely failed or eroded and has no load bearing
capacity. This method has been successfully applied to simulate high
complex phenomena such dynamic fragmentation of metals \cite{Li_2015},
nonetheless in quasi-brittle materials, simulations performed with
this approach exhibit an overestimation of tensile stress and the
strain peaks. To overcome this limitation, Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} introduced the concept of
eigensoftening. Engineered for quasi-brittle materials, this approach
is devoted to capture the gradual rather than abrupt dissipation of
the fracture energy. It is inspired in the concept of the crack band
model \cite{Bazant83}, since energy dissipation is thorough the
softened (or failed) volume. Due to the similarities between the OTM, 
where this procedure was developed, and the MPM, we propose

The eigensoftening algorithm together with a explicit predictor
corrector and the local \textit{max-ent} approximation technique
offers a new perspective in the MPM fracture simulation.

To overcome the difficulties of the e

In the present research we will introduce the

In the current work we attempt to model the fracture propagation
employing the MPM and introducing both eigenerosion and eigensoftening
\cite{Navas_2017_ES,Navas_2018_ES} and the concept of the
eigenfracture under this framework.

Even thought the method has been successfully applied to dynamic
fragmentation of metals, quantitative validations, such as stress or 
strain levels near or at the crack set compared to experimental
measures are still lacking. Having in mind that the eigenerosion
approach was intended for perfectly brittle fracture, an
eigensoftening concept is developed for quasi-brittle materials.

This document is structured as follows

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The meshfree methodology}
\label{sec:2}
The aim of this section is to provide an overview of the special
techniques employed to face the fracture problem under the MPM
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the explicit-predictor algorithm for the
MPM will be exposed, next the local \textit{max-ent} approximants are introduced in
\ref{sec:2.2} as an accurate alternative technique to interpolate data
between particles and nodes, and finally the fracture algorithms based
in the eigendeformations are presented in \ref{sec:2.3}.

\subsection{The Material Point Method}
\label{sec:2.1}
The MPM~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method
where particles carries on all the physical information and a set of
background nodes is employed to compute the equilibrium
equation. Since the MPM possesses the advantages of both Lagrangian
and Eulerian descriptions, no element distortion exists in the MPM,
therefore it is an appropriate and efficient method in solving
problems involving extremely large deformation and moving
discontinuities such fracture evolution. For the spatial
discretization, two sets of points are introduced in 
the MPM. First, the nodes. This points are considered fixed in the
space and are in charge of computing all the kinematic fields such
forces $f_I$, accelerations $a_I$ and velocities $v_I$. And second the
material points or particles. They are in charge of the discretization
of the continuum, and store the local state ($\sigma_p,
\varepsilon_p$). Without loosing generality, the MPM algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation thorough a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration
scheme is adopted. The purpose of this choice is motivated due its proved
robustness and stability in numerical calculations. In the first stage
(i), the nodal velocity predictor is computed following
\eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocity are
obtained, the essential boundary conditions are imposed. And in the
following, a Eulerian phase (ii) is computed in the set of nodes in a
FEM-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are computed thorough the
equilibrium equation. Next the nodal velocities are corrected in a
\textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}
\end{equation}
Finally updated the particles are advected in the Lagrangian stage (iii) using nodal values as,
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k} 
\end{align}
The complete pseudo-algorithm it is summarized in \ref{sec:expl-pred-corr}.

@@

\sum_I N_I |\vec{x}_p - \vec{x}_I |^2$ a magnitude of the shape
function width. The regularization o \textit{thermalization} parameter
between the two criterion $\beta$ has Pareto optimal values in the
range $(0,\infty)$. The unique solution of the local \textit{max-ent} problem
(LME)$_\beta$ is,

@@

a limit values of the influence radius and is used thereafter to

### 2020-04-25 _region_.tex
Cambios 24-4-2020

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}

The simulation of fracture propagation in a more accurate and
effective way can be considered as one of the original drivers for
developing novel spatial discretization methods such as meshfree
methods like the material point method (MPM). Presence of cracks are a
violation of the continuity requirement of the finite element method
(FEM). To overcome it, numerous of numerical artifacts has been
proposed with the aim of reproduce such a complex behaviour. This
techniques vary from employing cohesive approaches
\cite{Barenblatt,Hilleborg_1976}, by adaptively inserting cohesive
elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid elements
boundaries, or handling arbitrary cracks paths by level set
representation of the fracture surface \cite{Belytschko_03}.

The severe
limitations of the mesh-dependent methods to simulate fracture
process, have motivated the development of mesh-free technique to
reproduce crack propagation such the element-free Galerkin method
(EFGM) \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013},
the smoothed particle hydrodynamics (SPH) \cite{Wang_2020,Wang_2019},
the optimal transportation meshfree method (OTM)
\cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or peridynamics
\cite{HA_2011,RABCZUK_2017}.

By contrast, MPM does not suffer from the above difficulty since the
continuity requirement is less restrictive. Fracture can be described
in two ways. One is to remove the restriction of the single-valued
velocity field close to the crack by using two or more sets of nodes
\cite{Nairn_2006}. This method is based in to assign different labels
to distinguish if the material points and nodes are in the same side
of the crack or not. Under this approach the crack surface is
described with line segments in 2D and triangle patches in 3D
cases. This method was named as ``CRAcks with Material Points
(CRAMP)'' \cite{Nairn_2003}. In this method, the criteria for crack
propagation is based on such parameters as energy release rate
analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. The other approach is to use failed material
points to describe the crack evolution. In this method, the formation
of failed points describes the nucleation of cracks, and thereafter its propagation and
branching. Consequently, the position of the crack does not need to be
explicitly stated. This represents a significant advantage over the
``CRAMP''. Under this approach, the prediction the failure evolution
employing a decohesion model has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} or
Schreyer {\it et al.}\cite{Schreyer_2002}. And has been successfully
employed to simulate the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
(2004)\nocite{Sulsky_2004}.

\cite{Belytschko_04,Schmidt_2009,Pandolfi_2012}

The erosion of the material point means that each material point can
be either intact or be completely failed or eroded and has no load
bearing capacity, Even thought the method has been successfully
applied to dynamic fragmentation of metals, quantitative validations,
such as stress or strain levels near or at the crack set compared to
experimental measures are still lacking. Having in mind that the
eigenerosion approach was intended for perfectly brittle fracture, an
eigensoftening concept is developed for quasi-brittle materials

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The meshfree methodology}
\label{sec:2}

The aim of this section is to provide an overview of the special
techniques employed to face the fracture problem under the MPM
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the explicit-predictor algorithm for the
MPM will be exposed, next the local \textit{max-ent} approximants are introduced in
\ref{sec:2.2} as an accurate alternative technique to interpolate data
between particles and nodes, and finally the fracture algorithms based
in the eigendeformations are presented in \ref{sec:2.3}.

\subsection{The Material Point Method}
\label{sec:2.1}

The MPM~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method
where particles carries on all the physical information and a set of
background nodes is employed to compute the equilibrium
equation. Since the MPM possesses the advantages of both Lagrangian
and Eulerian descriptions, no element distortion exists in the MPM,
therefore it is an appropriate and efficient method in solving
problems involving extremely large deformation and moving
discontinuities such fracture evolution. For the spatial
discretization, two sets of points are introduced in 
the MPM. First, the nodes. This points are considered fixed in the
space and are in charge of computing all the kinematic fields such
forces $f_I$, accelerations $a_I$ and velocities $v_I$. And second the
material points or particles. They are in charge of the discretization
of the continuum, and store the local state ($\sigma_p,
\varepsilon_p$). Without loosing generality, the MPM algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation thorough a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration
scheme is adopted. The purpose of this choice is motivated due its proved
robustness and stability in numerical calculations. In the first stage
(i), the nodal velocity predictor is computed following
\eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocity are
obtained, the essential boundary conditions are imposed. And in the
following, a Eulerian phase (ii) is computed in the set of nodes in a
FEM-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are computed thorough the
equilibrium equation. Next the nodal velocities are corrected in a
\textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}
\end{equation}
Finally updated the particles are advected in the Lagrangian stage (iii) using nodal values as,
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k} 
\end{align}
The complete pseudo-algorithm it is summarized in \ref{sec:expl-pred-corr}.

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-660) }

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
Presence of cracks are a violation of the continuity requirement of
the finite element method (FEM). To overcome it, numerous of numerical
artifacts has been proposed with the aim of reproducing  such a
complex behaviour. This techniques vary from employing cohesive
approaches \cite{Barenblatt,Hilleborg_1976}, by adaptively inserting
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries, or handling arbitrary cracks paths by level set
representation of the fracture surface \cite{Belytschko_03}.

The simulation of fracture propagation in a more accurate and
effective way can be considered as one of the original drivers for
developing novel spatial discretization methods such as meshfree
techniques. Some examples of it are the material
point method (MPM) \cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
element-free Galerkin method (EFGM) \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013},
the smoothed particle hydrodynamics (SPH) \cite{Wang_2020,Wang_2019},
the optimal transportation meshfree method (OTM)
\cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or peridynamics
\cite{HA_2011,RABCZUK_2017} among others.

Fracture can be described in two ways. One is to remove the
restriction of the single-valued velocity field close to the crack by
using two or more sets of nodes \cite{Nairn_2006}. This method is
based in to assign different labels to distinguish if the material
points and nodes are in the same side of the crack or not. Under this
approach the crack surface is described with line segments in 2D and
triangle patches in 3D cases. This method was named as ``CRAcks with
Material Points (CRAMP)'' \cite{Nairn_2003}. In this method, the
criteria for crack propagation is based on such parameters as energy
release rate analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the
stress intensity factor or the J-integral discussed by Guo \& Nairn 
(2004)\cite{Nairn_2004}. The other approach is to use failed material
points to describe the crack evolution. In this method, the formation
of failed points describes the nucleation of cracks, and thereafter
its propagation and branching. Consequently, the position of the crack
does not need to be explicitly stated. This represents a significant
advantage over the ``CRAMP''. Under this approach, the prediction the
failure evolution employing a decohesion model has been discussed by
Chen {\it et al.}\cite{Zhenmao_2005} or Schreyer {\it et
  al.}\cite{Schreyer_2002}. And has been successfully employed to
simulate the fracture of brittle materials by Chen {\it et al.}
\cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}. Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
they approximate the crack set by means of eigen-deformations, wich
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle facture
was developed by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. In this technique the erosion of
the material point means that each material point can be either intact
or be completely failed or eroded and has no load bearing
capacity. This method has been successfully applied to simulate high
complex phenomena such dynamic fragmentation of metals \cite{Li_2015},
nonetheless in quasi-brittle materials, simulations performed with
this approach exhibit an overestimation of tensile stress and the
strain peaks. To overcome this limitation, Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} introduced the concept of
eigensoftening. Engineered for quasi-brittle materials, this approach
is devoted to capture the gradual rather than abrupt dissipation of
the fracture energy. It is inspired in the concept of the crack band
model \cite{Bazant83}, since energy dissipation is thorough the
softened (or failed) volume. Due to the similarities between the OTM, 
where this procedure was developed, and the MPM, we propose

To overcome the difficulties of the e

In the present research we will introduce the

In the current work we attempt to model the fracture propagation
employing the MPM and introducing both eigenerosion
 and eigensoftening \cite{Navas_2017_ES,Navas_2018_ES} and the concept of the
eigenfracture under this framework.

Even thought the method has been successfully applied to dynamic
fragmentation of metals, quantitative validations, such as stress or 
strain levels near or at the crack set compared to experimental
measures are still lacking. Having in mind that the eigenerosion
approach was intended for perfectly brittle fracture, an
eigensoftening concept is developed for quasi-brittle materials

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The meshfree methodology}
\label{sec:2}
The aim of this section is to provide an overview of the special
techniques employed to face the fracture problem under the MPM
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the explicit-predictor algorithm for the
MPM will be exposed, next the local \textit{max-ent} approximants are introduced in
\ref{sec:2.2} as an accurate alternative technique to interpolate data
between particles and nodes, and finally the fracture algorithms based
in the eigendeformations are presented in \ref{sec:2.3}.

\subsection{The Material Point Method}
\label{sec:2.1}
The MPM~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method
where particles carries on all the physical information and a set of
background nodes is employed to compute the equilibrium
equation. Since the MPM possesses the advantages of both Lagrangian
and Eulerian descriptions, no element distortion exists in the MPM,
therefore it is an appropriate and efficient method in solving
problems involving extremely large deformation and moving
discontinuities such fracture evolution. For the spatial
discretization, two sets of points are introduced in 
the MPM. First, the nodes. This points are considered fixed in the
space and are in charge of computing all the kinematic fields such
forces $f_I$, accelerations $a_I$ and velocities $v_I$. And second the
material points or particles. They are in charge of the discretization
of the continuum, and store the local state ($\sigma_p,
\varepsilon_p$). Without loosing generality, the MPM algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation thorough a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration
scheme is adopted. The purpose of this choice is motivated due its proved
robustness and stability in numerical calculations. In the first stage
(i), the nodal velocity predictor is computed following
\eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocity are
obtained, the essential boundary conditions are imposed. And in the
following, a Eulerian phase (ii) is computed in the set of nodes in a
FEM-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are computed thorough the
equilibrium equation. Next the nodal velocities are corrected in a
\textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}
\end{equation}
Finally updated the particles are advected in the Lagrangian stage (iii) using nodal values as,
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k} 
\end{align}
The complete pseudo-algorithm it is summarized in \ref{sec:expl-pred-corr}.

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-676) }
